---
---
title: "Hands-on Exercise 01: A Layered Grammar of Graphics: ggplot2 methods"
author: "Chen.PengWei"
date-modified: "last-modified"
execute:
  echo: true
  eval: true
  warning: false
  freeze: true
---

1 Learning outcome

In this chapter, you will learn the basic principles and essential components of ggplot2. At the same time, you will gain hands-on experience on using these components to plot statistical graphics based on the principle of Layered Grammar of Graphics.

2 Getting started

2.1 Installing and loading the required libraries

Before we get started, it is important for us to ensure that the required R packages have been installed.

::: callout-note
The code chunk on the right assumes that you already have `pacman` package installed.\
If not, please go ahead install `pacman` first.
:::

```{r}
exam_data <- read_csv("C:/Trista0114/ISSS608/hands-onEx/hands-onEx01/Exam_data.csv")
```

##Essential Grammatical Elements in ggplot2: data

```{r}
ggplot(data = exam_data)
```

##Essential Grammatical Elements in ggplot2: Aesthetic mappings

```{r}
ggplot(data=exam_data, 
       aes(x= MATHS))
```
