---
title: "Take-home Exercise 1: 
Creating data visualisation: Ship Performance in the Gulf of Guinea"
author: "Chen.PengWei"
date: "February 1, 2025"
date-modified: "last-modified"
execute:
  echo: true
  eval: true
  warning: false
  freeze: true
---

# 1. Overview

## 1.1 Background⛵**🌍**

::: {.callout-note style="background-color: #E8ECF1; border-left: 6px solid #90A4AE; padding: 15px; border-radius: 12px; box-shadow: 6px 6px 8px rgba(0,0,0,0.15);" appearance="simple" icon="false"}
Shipping plays an important role in global trade, facilitating the movement of goods across countries. It is crucial to understanding the profitability of different ship types and the relationship between profit and others variables. Various factors, such as operational costs, voyage revenue and fuel efficiency, may influence a ship’s profitability.

This study is a practice of data visualization based on the business perspective. Instead of building the model, this study will focus more on how to interpret insights and turn it into the visualization chart. The dataset used in this study is synthetically generated and is not sourced from actual shipping companies or vessels. The goal in this report is to find how different ship types perform in terms of profit, revenue and cost efficiency by using Exploratory Data Analysis (EDA).

The dataset be used is from Kaggle, ["Ship Performance Clustering Dataset"](https://www.kaggle.com/datasets/jeleeladekunlefijabi/ship-performance-clustering-dataset), provides various performance metrics, including operational costs, voyage revenue, speed, engine power, and fuel efficiency, enabling a comprehensive practice.
:::

## 1.2 Objective

The entire report is based on a business perspective, which means that all of the charts will focus on profitability. This take-home exercise is aim to do the following analysis by using Exploratory Data Analysis (EDA) methods and ggplot2 functions :

1️⃣ Profitability Analysis of Different Ship Types

-   Identify which ship type is the most profitable.

-   Analyze which ship type incurs the highest loss.

-   Analyze the investability in different ship type.

2️⃣ Key Influences on Profitability

-   Use Exploratory Data Analysis (EDA) to uncover relationships between profit and others metrics.

# 2. Getting Started

::: panel-tabset
## Installing Libraries

```{r}
pacman::p_load(ggplot2,dplyr,patchwork,tidyverse,tidymodels,ggdist,ggridges, ggthemes,colorspace,scales,ggiraph,ggstatsplot)
```

## Importing Data

The Ship Performance Dataset is a synthetic yet realistic collection of data designed to represent key operational metrics and attributes of various ship types in the Gulf of Guinea from Kaggle, ["Ship Performance Clustering Dataset"](https://www.kaggle.com/datasets/jeleeladekunlefijabi/ship-performance-clustering-dataset).

```{r}
ship_original <- read_csv("data/Ship_Performance_Dataset.csv")
```

## Understanding Data

The data set consists of 18 variables and 2736 observations. The variables including 12 numerical data and 6 Categorical data. We will do data filtering in the further step and most of the EDA will be associated with profit and ship type, which means not all the variables will be used to analyse in this case.

Numerical Variable

-   Speed_Over_Ground_knots: Average speed of the ship over water (in knots).

-   Engine_Power_kW: Engine power output (in kilowatts).

-   Distance_Traveled_nm: Total distance traveled by the ship (in nautical miles).

-   Operational_Cost_USD: Total operational cost per voyage (in USD).

-   Revenue_per_Voyage_USD: Revenue generated per voyage (in USD).

-   Efficiency_nm_per_kWh: Energy efficiency calculated in nautical miles per kilowatt-hour.

-   Draft_meters

-   Cargo_Weight_tons

-   Turnaround_Time_hours

-   Seasonal_Impact_Score

-   Weekly_Voyage_Count

-   Average_Load_Percentage

Categorical Variable

-   Ship_Type: Type of ship (e.g., Tanker, Container Ship, Fish Carrier, Bulk Carrier).

-   Route_Type: Shipping route type (e.g., Short-haul, Long-haul, Transoceanic).

-   Engine_Type: Type of engine (e.g., Diesel, Heavy Fuel Oil).

-   Maintenance_Status: Maintenance condition of the ship (e.g., Fair, Critical, Good).

-   Weather_Condition: Prevailing weather conditions during voyages (e.g., Calm, Moderate, Rough).

-   Date: From June 2023 to June 2024

```{r}
ship_original
```
:::

## 2.1 Prepare data

::: panel-tabset
## Filtering Data

Before diving into the profitability analysis of different ship types, we will first filter out the ship type that is "None" which has no meaning and do not contribute to the analysis. There are the total of 136 rows with "None" as the ship type, accounting for around 5% in this dataset.

```{r}
ship = ship_original %>% filter(Ship_Type != "None")
```

## Checking Missing Value

As the result from the code, there is no missing value in this dataset.

```{r}
colSums(is.na(ship))
```

## Creating variable : profit

By analysis the profitability, we need to caculate the variable "profit" by using Revenue_per_Voyage_USD - Operational_Cost_USD. After adding this variable, the updated dataset contains 2,600 rows and 19 variables, including the newly created "profit" column.

```{r}
ship <- ship %>%
  mutate(profit = Revenue_per_Voyage_USD - Operational_Cost_USD)
```

```{r}
head(ship,10)
```
:::

# 3. Exploratory Data Analysis

## 3.1 Distribution and box plot of Profit by ship type

The box plot and the distribution shows the profit across four different ship types: Bulk Carrier, Container Ship, Fish Carrier, and Tanker. The plot illustrates that all four ship types have a negative profit, which is an important point and will be discussed in further steps. Despite some variations, the overall profit distributions is similar across ship types, suggesting no big differences in profitability. The median and mean profit for each ship type remains positive and have just slightly differences.

::: panel-tabset
## The plot

```{r}
#| echo: false
ggplot(ship, aes(x = Ship_Type, y = profit)) +
  stat_halfeye(adjust = 0.5, 
               justification = -0.1,
               .width = 0,
               point_colour = NA,
               fill = "gray",
               alpha = 0.5) +  
  geom_boxplot(width = 0.1, 
               fill = "white", 
               color = "black") +
  stat_summary(fun = mean, 
               geom = "point", 
               size = 1, 
               color = "red", 
               fill = "red") +
  scale_y_continuous(labels = scales::comma) +  
  labs(title = "Profit Distribution by Ship Type",
       x = "Ship Type", 
       y = "Profit (USD)") +
  theme_minimal() + 
  theme(plot.title = element_text(hjust = 0.5))
```

## The code

```{r,eval=FALSE}
ggplot(ship, aes(x = Ship_Type, y = profit)) +
  stat_halfeye(adjust = 0.5, 
               justification = -0.1,
               .width = 0,
               point_colour = NA,
               fill = "gray",
               alpha = 0.5) +  
  geom_boxplot(width = 0.1, 
               fill = "white", 
               color = "black") +
  stat_summary(fun = mean, 
               geom = "point", 
               size = 1, 
               color = "red", 
               fill = "red") +
  scale_y_continuous(labels = scales::comma) +  
  labs(title = "Profit Distribution by Ship Type",
       x = "Ship Type", 
       y = "Profit (USD)") +
  theme_minimal() + 
  theme(plot.title = element_text(hjust = 0.5))
```
:::

## 3.2 Bar chart of Average Profit by ship type

In the overall point of view in this data, every ship type earn the money. While the differences in profitability among the ship types are not highly significant, Fish Carriers appear to be the most profitable, followed by Container Ships and Bulk Carriers, with Tankers having the lowest average profit. The gap between Fish Carriers and Tanker is around 20,000 USD. Hover to the chart and you can see the tooltip.

::: panel-tabset
## The plot

```{r}
#| echo: false
ship_colors <- c(
  "Tanker" = "#D3A29E", 
  "Container Ship" = "#88939E", 
  "Fish Carrier" = "#A3B18A", 
  "Bulk Carrier" = "#B5A7B2"
)

ship_avg_profit <- ship %>%
  group_by(Ship_Type) %>%
  summarise(Average_Profit = mean(profit, na.rm = TRUE))

y_min <- floor(min(ship_avg_profit$Average_Profit) / 1000) * 1000 
y_max <- ceiling(max(ship_avg_profit$Average_Profit) / 1000) * 1000  
y_breaks <- seq(y_min, y_max, by = 5000)  
y_lower_limit <- y_min - 5000  

tooltip <- function(y, accuracy = 1) {
  profit <- scales::number(y, accuracy = accuracy, big.mark = ",")
  paste("Average Profit: ", profit, " USD")
}


gg_point <- ggplot(data = ship_avg_profit, 
                   aes(x = Ship_Type, y = Average_Profit, fill = Ship_Type)) +
  geom_col_interactive(aes(tooltip = tooltip(Average_Profit), 
                           data_id = Ship_Type)) + 
  scale_fill_manual(values = ship_colors) +  
  scale_y_continuous(breaks = y_breaks, labels = scales::comma) +  
  coord_cartesian(ylim = c(y_lower_limit, y_max)) +  
  labs(title = "Average Profit by Ship Type", 
       x = "Ship Type", y = "Average Profit (USD)") +
  theme_minimal() +
  theme(
    axis.text.y = element_text(size = 10, angle = 0, hjust = 1),  
    axis.text.x = element_text(size = 12), 
    plot.title = element_text(hjust = 0.5, size = 14, face = "bold"),
    legend.position = "none" 
  )

girafe(ggobj = gg_point, width_svg = 8, height_svg = 8 * 0.618,
       options = list(
         opts_hover_inv(css = "opacity: 0.5;"),  
         opts_hover(css = "opacity: 1;")        
       ))
```

## The code

```{r,eval=FALSE}
ship_colors <- c(
  "Tanker" = "#D3A29E", 
  "Container Ship" = "#88939E", 
  "Fish Carrier" = "#A3B18A", 
  "Bulk Carrier" = "#B5A7B2"
)

ship_avg_profit <- ship %>%
  group_by(Ship_Type) %>%
  summarise(Average_Profit = mean(profit, na.rm = TRUE))

y_min <- floor(min(ship_avg_profit$Average_Profit) / 1000) * 1000 
y_max <- ceiling(max(ship_avg_profit$Average_Profit) / 1000) * 1000  
y_breaks <- seq(y_min, y_max, by = 5000)  
y_lower_limit <- y_min - 5000  

tooltip <- function(y, accuracy = 1) {
  profit <- scales::number(y, accuracy = accuracy, big.mark = ",")
  paste("Average Profit: ", profit, " USD")
}


gg_point <- ggplot(data = ship_avg_profit, 
                   aes(x = Ship_Type, y = Average_Profit, fill = Ship_Type)) +
  geom_col_interactive(aes(tooltip = tooltip(Average_Profit), 
                           data_id = Ship_Type)) + 
  scale_fill_manual(values = ship_colors) +  
  scale_y_continuous(breaks = y_breaks, labels = scales::comma) +  
  coord_cartesian(ylim = c(y_lower_limit, y_max)) +  
  labs(title = "Average Profit by Ship Type", 
       x = "Ship Type", y = "Average Profit (USD)") +
  theme_minimal() +
  theme(
    axis.text.y = element_text(size = 10, angle = 0, hjust = 1),  
    axis.text.x = element_text(size = 12), 
    plot.title = element_text(hjust = 0.5, size = 14, face = "bold"),
    legend.position = "none" 
  )

girafe(ggobj = gg_point, width_svg = 8, height_svg = 8 * 0.618,
       options = list(
         opts_hover_inv(css = "opacity: 0.3;"),  
         opts_hover(css = "opacity: 1;")        
       ))
```
:::

## 3.3 Percentage and the numbers of Negative Profit in ship type

The bar chart displays the number of negative profit voyages for each ship type. The ranking of ship types in terms of negative profit voyages remains consistent with previous analyses of the average profit in ship type. The stacked bar chart illustrates the distribution of profit bins for each ship type, showing the proportion of positive and negative profits within each category. Since the total number of ships in each category is relatively balanced, the percentage of negative profit voyages across ship types does not demonstrate a significant difference.

Overall, there is just a slightly difference in the number of negative profit in ship type, which highlight the idea that profitability challenges are somewhat evenly distributed across ship categories.

::: panel-tabset
## The plot

```{r}
#| echo: false
negative_profit_counts <- ship %>%
  filter(profit <= 0) %>%  
  group_by(Ship_Type) %>%
  summarise(
    Negative_Profit_Count = n(),
    Negative_Profit_Percentage = Negative_Profit_Count / sum(Negative_Profit_Count) * 100, 
    .groups = "drop"
  )

lower_limit <- 50
upper_limit <- max(negative_profit_counts$Negative_Profit_Count) * 1.1 


p1 <- ggplot(data = negative_profit_counts, 
             aes(x = Ship_Type, y = Negative_Profit_Count, fill = Ship_Type)) +
  geom_col_interactive(
    aes(tooltip = paste0("Number of Negative and Zero Profit Voyages: ", scales::comma(Negative_Profit_Count))),
    show.legend = FALSE
  ) +
  geom_text(aes(label = scales::comma(Negative_Profit_Count)),  
            vjust = -0.5, size = 3) +
  scale_fill_manual(values = ship_colors) +
  scale_y_continuous(labels = scales::comma) +  
  coord_cartesian(ylim = c(lower_limit, upper_limit)) +  
  labs(
    title = "Number of Negative and Zero Profit Voyages by Ship Type",
    x = "Ship Type",
    y = "Number of Voyages with Negative and Zero Profit"
  ) +
  theme_minimal() +
  theme(
    plot.title = element_text(hjust = 0.5, size = 8, face = "bold"),
    axis.text.x = element_text(size = 6, angle = 0, hjust = 0.5),
    axis.text.y = element_text(size = 7),
    axis.title.x = element_text(size = 8),
    axis.title.y = element_text(size = 8)
  )

ship_bin <- ship %>%
  mutate(profit_bins = cut(profit,
                           breaks = c(-Inf, 0, Inf),  
                           labels = c("Negative", "Positive"),  
                           include.lowest = TRUE))  

ship_bin <- ship_bin %>% filter(!is.na(profit_bins))


p2 <- ggbarstats(
  data = ship_bin,
  x = profit_bins,
  y = Ship_Type,
  title = "Distribution of Profit Bins by Ship Type",
  results.subtitle = FALSE,
  ggtheme = ggplot2::theme_minimal()
) +
  ggplot2::scale_y_continuous(labels = scales::percent_format(scale = 1)) +
  ggplot2::scale_fill_manual(values = c("Negative" = "salmon", "Positive" = "lightblue")) + 
  ggplot2::theme(
    plot.title = ggplot2::element_text(hjust = 0.5, size = 9,face = "bold"), 
    axis.title.x = ggplot2::element_text(size = 9),  
    axis.title.y = ggplot2::element_text(size = 9),  
    axis.text.x = ggplot2::element_text(size = 6,hjust = 0.5),  
    axis.text.y = ggplot2::element_text(size = 8),  
    legend.text = ggplot2::element_text(size = 9),  
    legend.title = ggplot2::element_text(size = 9)  
  )

p1 + p2
```

## The code

```{r,eval=FALSE}
negative_profit_counts <- ship %>%
  filter(profit <= 0) %>%  
  group_by(Ship_Type) %>%
  summarise(
    Negative_Profit_Count = n(),
    Negative_Profit_Percentage = Negative_Profit_Count / sum(Negative_Profit_Count) * 100, 
    .groups = "drop"
  )

lower_limit <- 50
upper_limit <- max(negative_profit_counts$Negative_Profit_Count) * 1.1 


p1 <- ggplot(data = negative_profit_counts, 
             aes(x = Ship_Type, y = Negative_Profit_Count, fill = Ship_Type)) +
  geom_col_interactive(
    aes(tooltip = paste0("Number of Negative and Zero Profit Voyages: ", scales::comma(Negative_Profit_Count))),
    show.legend = FALSE
  ) +
  geom_text(aes(label = scales::comma(Negative_Profit_Count)),  
            vjust = -0.5, size = 3) +
  scale_fill_manual(values = ship_colors) +
  scale_y_continuous(labels = scales::comma) +  
  coord_cartesian(ylim = c(lower_limit, upper_limit)) +  
  labs(
    title = "Number of Negative and Zero Profit Voyages by Ship Type",
    x = "Ship Type",
    y = "Number of Voyages with Negative and Zero Profit"
  ) +
  theme_minimal() +
  theme(
    plot.title = element_text(hjust = 0.5, size = 6, face = "bold"),
    axis.text.x = element_text(size = 6, angle = 0, hjust = 0.5),
    axis.text.y = element_text(size = 7),
    axis.title.x = element_text(size = 8),
    axis.title.y = element_text(size = 8)
  )

ship_bin <- ship %>%
  mutate(profit_bins = cut(profit,
                           breaks = c(-Inf, 0, Inf),  
                           labels = c("Negative", "Positive"),  
                           include.lowest = TRUE))  

ship_bin <- ship_bin %>% filter(!is.na(profit_bins))


p2 <- ggbarstats(
  data = ship_bin,
  x = profit_bins,
  y = Ship_Type,
  title = "Distribution of Profit Bins by Ship Type",
  results.subtitle = FALSE,
  ggtheme = ggplot2::theme_minimal()
) +
  ggplot2::scale_y_continuous(labels = scales::percent_format(scale = 1)) +
  ggplot2::scale_fill_manual(values = c("Negative" = "salmon", "Positive" = "lightblue")) + 
  ggplot2::theme(
    plot.title = ggplot2::element_text(hjust = 0.5, size = 9,face = "bold"), 
    axis.title.x = ggplot2::element_text(size = 9),  
    axis.title.y = ggplot2::element_text(size = 9),  
    axis.text.x = ggplot2::element_text(size = 6,hjust = 0.5),  
    axis.text.y = ggplot2::element_text(size = 8),  
    legend.text = ggplot2::element_text(size = 9),  
    legend.title = ggplot2::element_text(size = 9)  
  )

p1 + p2
```
:::

## 3.4 Distrubution of Negative and Zero Profit in ship type

Tanker has the highest number of negative profit voyages in the previous plot but the lowest average loss in this box plot. In contrast, Bulk Carrier has fewer loss-making voyages but the highest average loss, indicating larger financial risks per voyage, with the gap around 30,000 USD compared to Tanker.

The Density Plot further prove this. Tanker’s distribution is closer to zero, reflecting smaller losses, while Bulk Carrier skews left, showing higher loss .

This suggests that Tankers experience frequent but moderate losses, while Bulk Carriers face larger but less frequent losses, highlighting different risk profiles for each ship type.

::: panel-tabset
## The plot

```{r}
#| echo: false

negative_profit_ship <- ship %>%
  filter(profit <= 0) %>%  
  mutate(Ship_Type = factor(Ship_Type, 
                            levels = rev(c("Bulk Carrier", "Fish Carrier", "Container Ship", "Tanker"))))


p3 <- ggplot(negative_profit_ship, 
       aes(x = profit, y = Ship_Type, fill = Ship_Type)) +
  geom_density_ridges(scale = 5, alpha = 0.8, rel_min_height = 0.01) +  
  scale_fill_manual(values = ship_colors) +  
  scale_x_continuous(labels = scales::comma) + 
  labs(
    title = "Density Plot of Negative Profit by Ship Type",
    x = "Profit (USD)",
    y = "Ship Type"
  ) +
  theme_minimal() +
  theme(
    plot.title = element_text(hjust = 0.5, size = 9, face = "bold"),
    axis.title.x = element_text(size = 12),
    axis.title.y = element_text(size = 12),
    legend.position = "none"  
  )


mean_values <- negative_profit_ship %>%
  group_by(Ship_Type) %>%
  summarise(mean_profit = mean(profit), .groups = "drop")

p4 <- ggplot(negative_profit_ship, aes(x = Ship_Type, y = profit, fill = Ship_Type)) +
  geom_boxplot(alpha = 0.8, outlier.color = "black", outlier.shape = 16) +  
  scale_fill_manual(values = ship_colors) +  
  scale_y_continuous(labels = scales::comma) +  
  labs(
    title = "Box Plot of Negative Profit by Ship Type",
    x = "Ship Type",
    y = "Profit (USD)"
  ) +

  geom_point(data = mean_values, 
             aes(x = Ship_Type, y = mean_profit), 
             color = "red", 
             size = 2) +

  geom_text(data = mean_values, 
            aes(x = Ship_Type, y = mean_profit, 
                label = paste0(scales::comma(round(mean_profit)))),
            vjust = 1.5, size = 3, color = "black") + 
  theme_minimal() +
  theme(
    plot.title = element_text(hjust = 0.5, size = 9, face = "bold"),
    axis.title.x = element_text(size = 12),
    axis.title.y = element_text(size = 12),
    axis.text.x = ggplot2::element_text(size = 6,hjust = 0.5),
    legend.position = "none"
  )

p4 + p3
```

## The code

```{r,eval=FALSE}
negative_profit_ship <- ship %>%
  filter(profit <= 0) %>%  
  mutate(Ship_Type = factor(Ship_Type, 
                            levels = rev(c("Bulk Carrier", "Fish Carrier", "Container Ship", "Tanker"))))


p3 <- ggplot(negative_profit_ship, 
       aes(x = profit, y = Ship_Type, fill = Ship_Type)) +
  geom_density_ridges(scale = 5, alpha = 0.8, rel_min_height = 0.01) +  
  scale_fill_manual(values = ship_colors) +  
  scale_x_continuous(labels = scales::comma) + 
  labs(
    title = "Density Plot of Negative Profit by Ship Type",
    x = "Profit (USD)",
    y = "Ship Type"
  ) +
  theme_minimal() +
  theme(
    plot.title = element_text(hjust = 0.5, size = 9, face = "bold"),
    axis.title.x = element_text(size = 12),
    axis.title.y = element_text(size = 12),
    legend.position = "none"  
  )


mean_values <- negative_profit_ship %>%
  group_by(Ship_Type) %>%
  summarise(mean_profit = mean(profit), .groups = "drop")

p4 <- ggplot(negative_profit_ship, aes(x = Ship_Type, y = profit, fill = Ship_Type)) +
  geom_boxplot(alpha = 0.8, outlier.color = "black", outlier.shape = 16) +  
  scale_fill_manual(values = ship_colors) +  
  scale_y_continuous(labels = scales::comma) +  
  labs(
    title = "Box Plot of Negative Profit by Ship Type",
    x = "Ship Type",
    y = "Profit (USD)"
  ) +

  geom_point(data = mean_values, 
             aes(x = Ship_Type, y = mean_profit), 
             color = "red", 
             size = 2) +

  geom_text(data = mean_values, 
            aes(x = Ship_Type, y = mean_profit, 
                label = paste0(scales::comma(round(mean_profit)))),
            vjust = 1.5, size = 3, color = "black") + 
  theme_minimal() +
  theme(
    plot.title = element_text(hjust = 0.5, size = 9, face = "bold"),
    axis.title.x = element_text(size = 12),
    axis.title.y = element_text(size = 12),
    axis.text.x = ggplot2::element_text(size = 6,hjust = 0.5),
    legend.position = "none"
  )

p4 + p3
```
:::

## 3.5 Distrubution of Positive Profit in ship type

Based on plots , Bulk Carrier appears to be a high-risk ship type. While it has the potential for higher profits when successful, its advantage over other ship types in positive profits is not as significant as its disadvantage in negative profits. Fish Carrier, on the other hand, maintains a relatively stable profit profile, with a balanced performance in both positive and negative profit scenarios. Selecting Fish Carrier as an investment or operational focus could be a more stable choice.

::: callout-caution
## 💰However..

The reasons behind the negative profits of container ship and bulk carriers are worth exploring. Identifying and addressing these factors can help decrease financial risk and may become a profitable option due to the higher returns in a positive profit .
:::

::: panel-tabset
## The plot

```{r}
#| echo: false
Positive_profit_ship <- ship %>%
  filter(profit > 0) %>%  
  mutate(Ship_Type = factor(Ship_Type, 
                            levels = rev(c("Bulk Carrier", "Fish Carrier", "Container Ship", "Tanker"))))


p5 <- ggplot(Positive_profit_ship, 
       aes(x = profit, y = Ship_Type, fill = Ship_Type)) +
  geom_density_ridges(scale = 5, alpha = 0.8, rel_min_height = 0.01) +  
  scale_fill_manual(values = ship_colors) +  
  scale_x_continuous(labels = scales::comma) + 
  labs(
    title = "Density Plot of Positive Profit by Ship Type",
    x = "Profit (USD)",
    y = "Ship Type"
  ) +
  theme_minimal() +
  theme(
    plot.title = element_text(hjust = 0.5, size = 9, face = "bold"),
    axis.title.x = element_text(size = 12),
    axis.title.y = element_text(size = 12),
    legend.position = "none"  
  )


mean_values <- Positive_profit_ship %>%
  group_by(Ship_Type) %>%
  summarise(mean_profit = mean(profit), .groups = "drop")

p6 <- ggplot(Positive_profit_ship, aes(x = Ship_Type, y = profit, fill = Ship_Type)) +
  geom_boxplot(alpha = 0.8, outlier.color = "black", outlier.shape = 16) +  
  scale_fill_manual(values = ship_colors) +  
  scale_y_continuous(labels = scales::comma) +  
  labs(
    title = "Box Plot of Postive  Profit by Ship Type",
    x = "Ship Type",
    y = "Profit (USD)"
  ) +

  geom_point(data = mean_values, 
             aes(x = Ship_Type, y = mean_profit), 
             color = "red", 
             size = 2) +

  geom_text(data = mean_values, 
            aes(x = Ship_Type, y = mean_profit, 
                label = paste0(scales::comma(round(mean_profit)))),
            vjust = -0.2, size = 3, color = "black") + 
  theme_minimal() +
  theme(
    plot.title = element_text(hjust = 0.5, size = 9, face = "bold"),
    axis.title.x = element_text(size = 12),
    axis.title.y = element_text(size = 12),
    axis.text.x = ggplot2::element_text(size = 6,hjust = 0.5),
    legend.position = "none"
  )

p6 + p5
```

## The code

```{r,eval=FALSE}
Positive_profit_ship <- ship %>%
  filter(profit > 0) %>%  
  mutate(Ship_Type = factor(Ship_Type, 
                            levels = rev(c("Bulk Carrier", "Fish Carrier", "Container Ship", "Tanker"))))


p5 <- ggplot(Positive_profit_ship, 
       aes(x = profit, y = Ship_Type, fill = Ship_Type)) +
  geom_density_ridges(scale = 5, alpha = 0.8, rel_min_height = 0.01) +  
  scale_fill_manual(values = ship_colors) +  
  scale_x_continuous(labels = scales::comma) + 
  labs(
    title = "Density Plot of Positive Profit by Ship Type",
    x = "Profit (USD)",
    y = "Ship Type"
  ) +
  theme_minimal() +
  theme(
    plot.title = element_text(hjust = 0.5, size = 9, face = "bold"),
    axis.title.x = element_text(size = 12),
    axis.title.y = element_text(size = 12),
    legend.position = "none"  
  )


mean_values <- Positive_profit_ship %>%
  group_by(Ship_Type) %>%
  summarise(mean_profit = mean(profit), .groups = "drop")

p6 <- ggplot(Positive_profit_ship, aes(x = Ship_Type, y = profit, fill = Ship_Type)) +
  geom_boxplot(alpha = 0.8, outlier.color = "black", outlier.shape = 16) +  
  scale_fill_manual(values = ship_colors) +  
  scale_y_continuous(labels = scales::comma) +  
  labs(
    title = "Box Plot of Postive  Profit by Ship Type",
    x = "Ship Type",
    y = "Profit (USD)"
  ) +

  geom_point(data = mean_values, 
             aes(x = Ship_Type, y = mean_profit), 
             color = "red", 
             size = 2) +

  geom_text(data = mean_values, 
            aes(x = Ship_Type, y = mean_profit, 
                label = paste0(scales::comma(round(mean_profit)))),
            vjust = -0.2, size = 3, color = "black") + 
  theme_minimal() +
  theme(
    plot.title = element_text(hjust = 0.5, size = 9, face = "bold"),
    axis.title.x = element_text(size = 12),
    axis.title.y = element_text(size = 12),
    axis.text.x = ggplot2::element_text(size = 6,hjust = 0.5),
    legend.position = "none"
  )

p6 + p5
```
:::

## 3.6 Relationship between Revenue and Cost

In a general business case, revenue and cost usually show a positive correlation, because higher operating costs often mean larger business revenue scale. However, in this data, the Pearson correlation coefficient between the two ( 𝑟 r) is only 0.02, and the p-value is 0.22, showing that there is no significant linear relationship between the two.

::: callout-note
## This may mean 🔍🤔

-   Business model impact: Some ship types or businesses may have a fixed revenue structure, and even if costs increase, revenue will not increase accordingly.
-   Cost control issues: If the operating efficiency of some ships is low, even if higher costs are invested, revenue cannot be effectively increased.
-   Data issues: There may be other variables affecting revenue, such as market demand, freight rates, etc., rather than being influenced by costs. Another reason may because this synthetically generated dataset does not align with real-world scenario.
:::

::: panel-tabset
## The plot

```{r}
#| echo: false
ggscatterstats(
  data = ship,
  x = Operational_Cost_USD,
  y = Revenue_per_Voyage_USD,
  marginal = FALSE,
  )
```

## The code

```{r,eval=FALSE}
ggscatterstats(
  data = ship,
  x = Operational_Cost_USD,
  y = Revenue_per_Voyage_USD,
  marginal = FALSE,
  )
```
:::

## 3.7 Relationship between Efficiency and Profit

The scatter plot shows that there is no significant relationship between energy efficiency and profit , while higher energy efficiency should bring better economic benefits generally from a business perspective. The expected correlation does not exist in this data set.

::: callout-note
## This may mean 🔍🤔

-   Other factors (such as freight volume, route selection, freight market fluctuations) may have a greater impact on profits than pure energy efficiency metrics.
-   High energy efficiency does not necessarily helpful to the profit: Even if the ship's energy efficiency is high, profits may not be improved due to other maintenance costs, rent, labor and other expenses.
:::

::: panel-tabset
## The plot

```{r}
#| echo: false
ggscatterstats(
  data = ship,
  x = Efficiency_nm_per_kWh,
  y = profit,
  marginal = FALSE,
  )
```

## The code

```{r,eval=FALSE}
ggscatterstats(
  data = ship,
  x = Efficiency_nm_per_kWh,
  y = profit,
  marginal = FALSE,
  )
```
:::

# 4. Conclusion🚢💡

Fish Carrier are the most balanced ship type in terms of profitability and risk, while bulk carriers offer high returns but come with significant financial risk. It is worth to develop into the reasons behind this, as Bulk Carriers and Container Ship have the potential to become the profitable ship type if their financial risk can be decreased.

Additionally, revenue and cost do not show a significant correlation, meaning that increasing operational costs does not necessarily lead to higher revenue. This contradicts basic business principles and may suggest that other factors influence revenue in this industry or this synthetically generated dataset does not align with real-world scenario.

# 5. Reference

 [Kam, T.S. (2025). Visual Statistical Analysis.](https://r4va.netlify.app/chap10#getting-started-1)
