---
title: "Hands on Exercise 08-3: Analytical Mapping"
author: "Chen.PengWei"
date-modified: "last-modified"
execute:
  echo: true
  eval: true
  warning: false
  freeze: true
---

# 1. Overview

In this in-class exercise, you will learn:

-   Creating rate map
-   Creating percentile map
-   Creating boxmap

# 2. Getting Started

::: panel-tabset
## Installing Libraries

install and load **sf**, **tmap** and **tidyverse** packages

```{r}
pacman::p_load(tmap, tidyverse, sf)
```

## Importing Data

The data set is a polygon feature data.frame providing information on water point of Nigeria at the LGA level.

```{r}
NGA_wp <- read_rds("data/rds/NGA_wp.rds")
```
:::

# 3. Basic Choropleth Mapping

## 3.1 Visualising distribution 

Plot a choropleth map showing the distribution of non-function water point by LGA

```{r}
p1 <- tm_shape(NGA_wp) +
  tm_fill("wp_functional",
          n = 10,
          style = "equal",
          palette = "Blues") +
  tm_borders(lwd = 0.1,
             alpha = 1) +
  tm_layout(main.title = "Distribution of functional water point by LGAs",
            legend.outside = FALSE)

p2 <- tm_shape(NGA_wp) +
  tm_fill("total_wp",
          n = 10,
          style = "equal",
          palette = "Blues") +
  tm_borders(lwd = 0.1,
             alpha = 1) +
  tm_layout(main.title = "Distribution of total  water point by LGAs",
            legend.outside = FALSE)

tmap_arrange(p2, p1, nrow = 1)
```

# 🛠️4. Choropleth Map for Rates

In much of our readings we have now seen the importance to map rates rather than counts of things, and that is for the simple reason that water points are not equally distributed in space. That means that if we do not account for how many water points are somewhere, we end up mapping **total water point** size rather than our topic of interest.

We will tabulate the proportion of functional water points and the proportion of non-functional water points in each LGA. In the following code chunk, `mutate(`) from **dplyr** package is used to derive two fields, namely *pct_functional* and *pct_nonfunctional*.

Plot a choropleth map showing the distribution of percentage functional water point by LGA

```{r}
NGA_wp <- NGA_wp %>%
  mutate(pct_functional = wp_functional/total_wp) %>%
  mutate(pct_nonfunctional = wp_nonfunctional/total_wp)
NGA_wp
```

```{r}
tm_shape(NGA_wp) +
  tm_fill("pct_functional",
          n = 10,
          style = "equal",
          palette = "Blues",
          legend.hist = TRUE) +
  tm_borders(lwd = 0.1,
             alpha = 1) +
  tm_layout(main.title = "Rate map of functional water point by LGAs",
            legend.outside = TRUE)
```

# 5. Extreme Value Maps

**Extreme value maps are variations of common choropleth maps where the classification is designed to highlight extreme values at the lower and upper end of the scale**, with the goal of identifying outliers. These maps were developed in the spirit of spatializing EDA, i.e., adding spatial features to commonly used approaches in non-spatial EDA (Anselin 1994).

## 5.1 Percentile Map

The percentile map is a special type of quantile map with six specific categories: 0-1%,1-10%, 10-50%,50-90%,90-99%, and 99-100%. 

### 5.1.1Data Preparation

:::: panel-tabset
## step1: Exclude records with NA

```{r}
NGA_wp <- NGA_wp %>%
  drop_na()
```

## step2: Creating customised classification and extracting values

st_set_geometry(NULL): Removes the geometry information and ensures that var only contains values.\
var\[,1\]: Extract the first column of var (because var is a data.frame or tibble, it needs to be valued this way). quantile(..., percent): Calculate the value of var\[,1\] at the percentile defined by percent

```{r}
percent <- c(0,.01,.1,.5,.9,.99,1)
var <- NGA_wp["pct_functional"] %>%
  st_set_geometry(NULL)
quantile(var[,1], percent)
```

::: callout-important
When variables are extracted from an sf data.frame, the geometry is extracted as well. For mapping and spatial manipulation, this is the expected behavior, but many base R functions cannot deal with the geometry. Specifically, the `quantile()` gives an error. As a result `st_set_geomtry(NULL)` is used to drop geomtry field.
:::

## Why writing functions?

Writing a function has three big advantages over using copy-and-paste:

-   You can give a function an evocative name that makes your code easier to understand.
-   As requirements change, you only need to update code in one place, instead of many.
-   You eliminate the chance of making incidental mistakes when you copy and paste (i.e. updating a variable name in one place, but not in another).
::::

Firstly, we will write an R function as shown below to extract a variable (i.e. *wp_nonfunctional*) as a vector out of an sf data.frame.

-   arguments:

    -   vname: variable name (as character, in quotes)
    -   df: name of sf data frame

-   returns:

    -   v: vector with values (without a column name)

✅ Extract single variable from sf data frame ✅ Remove geographical information and turn variables into ordinary vectors ✅ Return pure vector, no column names

```{r}
get.var <- function(vname,df) {
  v <- df[vname] %>% 
    st_set_geometry(NULL)
  v <- unname(v[,1])
  return(v)
}
```

### 5.1.2 percentile mapping

```{r}
percentmap <- function(vnam, df, legtitle=NA, mtitle="Percentile Map"){
  percent <- c(0,.01,.1,.5,.9,.99,1)
  var <- get.var(vnam, df)
  bperc <- quantile(var, percent)
  tm_shape(df) +
  tm_polygons() +
  tm_shape(df) +
     tm_fill(vnam,
             title=legtitle,
             breaks=bperc,
             palette="Blues",
          labels=c("< 1%", "1% - 10%", "10% - 50%", "50% - 90%", "90% - 99%", "> 99%"))  +
  tm_borders() +
  tm_layout(main.title = mtitle, 
            title.position = c("right","bottom"))
}

percentmap("total_wp", NGA_wp)
```

## 5.2 Box map

```{r}
ggplot(data = NGA_wp,
       aes(x = "",
           y = wp_nonfunctional)) +
  geom_boxplot()
```

-   Displaying summary statistics on a choropleth map by using the basic principles of boxplot.
-   To create a box map, a custom breaks specification will be used. However, there is a complication. The break points for the box map vary depending on whether lower or upper outliers are present.

The code chunk below is an R function that creating break points for a box map.

-   arguments:

    -   v: vector with observations
    -   mult: multiplier for IQR (default 1.5)

-   returns:

    -   bb: vector with 7 break points compute quartile and fences

bb is a numeric vector of length 7, used to store: \
bb\[1\]: lower bound (minimum value or lofence) \
bb\[2\]: first critical value (qv\[1\] or lofence) \
bb\[3\]: first quartile Q1 \
bb\[4\]: median Q2 \
bb\[5\]: third quartile Q3 \
bb\[6\]: second critical value (qv\[5\] or upfence) \
bb\[7\]: upper bound (maximum value or upfence)

If lofence is smaller than qv\[1\], which means there is no outlier, then lofence is directly set to bb\[1\]. Otherwise, set qv\[1\] to bb\[1\] and lofence to bb\[2\].

```{r}
boxbreaks <- function(v,mult=1.5) {
  qv <- unname(quantile(v))
  iqr <- qv[4] - qv[2]
  upfence <- qv[4] + mult * iqr
  lofence <- qv[2] - mult * iqr
  # initialize break points vector
  bb <- vector(mode="numeric",length=7)
  # logic for lower and upper fences
  if (lofence < qv[1]) {  # no lower outliers
    bb[1] <- lofence
    bb[2] <- floor(qv[1])
  } else {
    bb[2] <- lofence
    bb[1] <- qv[1]
  }
  if (upfence > qv[5]) { # no upper outliers
    bb[7] <- upfence
    bb[6] <- ceiling(qv[5])
  } else {
    bb[6] <- upfence
    bb[7] <- qv[5]
  }
  bb[3:5] <- qv[2:4]
  return(bb)
}
```

The code chunk below is an R function to extract a variable as a vector out of an sf data frame.

-   arguments:

    -   vname: variable name (as character, in quotes)

    -   df: name of sf data frame

-   returns:

    -   v: vector with values (without a column name)

```{r}
get.var <- function(vname,df) {
  v <- df[vname] %>% st_set_geometry(NULL)
  v <- unname(v[,1])
  return(v)
}
```

Let’s test the newly created function

```{r}
var <- get.var("wp_nonfunctional", NGA_wp) 
boxbreaks(var)
```

finally, we can create a box map.\
\
vnam: variable name (data column name, string format). \
df: Data frame of sf type, containing geospatial data. \
legtitle (optional): legend title (default is NA). \
mtitle (optional): map title (default is "Box Map"). \
mult (optional): IQR (interquartile range) multiples used to calculate outliers (default is 1.5).\
\
get.var(vnam, df): Extract the variable vnam from df of type sf and convert it into a pure vector (excluding geographical information). \
boxbreaks(var): Calculate 7 break points of this variable for Box Map visualization

```{r}
boxmap <- function(vnam, df, 
                   legtitle=NA,
                   mtitle="Box Map",
                   mult=1.5){
  var <- get.var(vnam,df)
  bb <- boxbreaks(var)
  tm_shape(df) +
    tm_polygons() +
  tm_shape(df) +
     tm_fill(vnam,title=legtitle,
             breaks=bb,
             palette="Blues",
          labels = c("lower outlier", 
                     "< 25%", 
                     "25% - 50%", 
                     "50% - 75%",
                     "> 75%", 
                     "upper outlier"))  +
  tm_borders() +
  tm_layout(main.title = mtitle, 
            title.position = c("left",
                               "top"))
}
tmap_mode("plot")
boxmap("wp_nonfunctional", NGA_wp)
```
