---
title: "Hands on Exercise 05-5: Treemap Visualisation with R"
author: "Chen.PengWei"
date-modified: "last-modified"
execute:
  echo: true
  eval: true
  warning: false
  freeze: true
---

# 1. Overview

------------------------------------------------------------------------

In this hands-on exercise, you will gain hands-on experiences on designing treemap using appropriate R packages. 

::: callout-note
## Things you will learn

-   2 Data preparation for treemap

-   3 Designing Treemap with **treemap** Package

    -   changing basic default : color / type

    -   *3.7 3.8 working with algorithm* argument/sortID

-   4 Designing Treemap with **treemapify** Package

-   5 Designing Interactive Treemap using **d3treeR**
:::

# 2. Getting Started

::: panel-tabset
## Installing Libraries

Before we get started, you are required to check if **treemap** and **tidyverse** pacakges have been installed in you R.

```{r}
pacman::p_load(treemap, treemapify, tidyverse) 
```

## Importing Data

In this exercise, *REALIS2018.csv* data will be used. This dataset provides information of private property transaction records in 2018. The dataset is extracted from REALIS portal (https://spring.ura.gov.sg/lad/ore/login/index.cfm) of Urban Redevelopment Authority (URA).

```{r}
realis2018 <- read_csv("data/realis2018.csv")
```
:::

:::: panel-tabset
## Data Wrangling and Manipulation

The data.frame *realis2018* is in trasaction record form, which is highly disaggregated and not appropriate to be used to plot a treemap. In this section, we will perform the following steps to manipulate and prepare a data.frtame that is appropriate for treemap visualisation:

-   group transaction records by *Project Name*, *Planning Region*, *Planning Area*, *Property Type* and *Type of Sale*, and

-   compute *Total Unit Sold*, *Total Area*, *Median Unit Price* and *Median Transacted Price* by applying appropriate summary statistics on *No. of Units*, *Area (sqm)*, *Unit Price (\$ psm)* and *Transacted Price (\$)* respectively.

Two key verbs of **dplyr** package, namely: *group_by()* and *summarize()* will be used to perform these steps.

::: callout-note
## Grouping affects

-   `select()`

    -   Same as ungrouped, but grouping variables are automatically retained.

-   `arrange()`

    -   The default is the same as when not grouped.

    -   Setting .`by_group = TRUE` will sort by the group variable first.

-   mutate() & filter()

    -   Most useful when paired with window functions (such as `rank()`, `min(x) == x`).

    -   Applies to each set of internal calculations.

-   `sample_n()` & `sample_frac()`

    -   A specified amount `(sample_n())` or proportion (`sample_frac()`) of data is extracted within each group. s

-   `summarize()`

    -   Calculate summary statistics for each set of data, such as mean(), sum(), n(), etc.
:::

Students who are new to **dplyr** methods should consult [Introduction to dplyr](https://cran.r-project.org/web/packages/dplyr/vignettes/dplyr.html) before moving on to the next section.

## Grouped summaries without the Pipe

Aggregation functions such as *sum()* and *meadian()* obey the usual rule of missing values: if there’s any missing value in the input, the output will be a missing value. The argument *na.rm = TRUE* removes the missing values prior to computation.

```{r}
realis2018_grouped <- group_by(realis2018, `Project Name`,
                               `Planning Region`, `Planning Area`, 
                               `Property Type`, `Type of Sale`)
realis2018_summarised <- summarise(realis2018_grouped, 
                          `Total Unit Sold` = sum(`No. of Units`, na.rm = TRUE),
                          `Total Area` = sum(`Area (sqm)`, na.rm = TRUE),
                          `Median Unit Price ($ psm)` = median(`Unit Price ($ psm)`, na.rm = TRUE), 
                          `Median Transacted Price` = median(`Transacted Price ($)`, na.rm = TRUE))
```

```{r}
head(realis2018_summarised,10)
```

## Grouped summaries with the pipe

The code chunk below shows a more efficient way to tackle the same processes by using the *pipe*, %\>%:

```{r}
realis2018_summarised <- realis2018 %>% 
  group_by(`Project Name`,`Planning Region`, 
           `Planning Area`, `Property Type`, 
           `Type of Sale`) %>%
  summarise(`Total Unit Sold` = sum(`No. of Units`, na.rm = TRUE), 
            `Total Area` = sum(`Area (sqm)`, na.rm = TRUE),
            `Median Unit Price ($ psm)` = median(`Unit Price ($ psm)`, na.rm = TRUE),
            `Median Transacted Price` = median(`Transacted Price ($)`, na.rm = TRUE))
```
::::

# 3. Designing Treemap with treemap Package

**treemap** package is a R package specially designed to offer great flexibility in drawing treemaps. The core function, namely: *treemap()* offers at least 43 arguments. In this section, we will only explore the major arguments for designing elegent and yet truthful treemaps.

## 3.1 Designing a static treemap

In this section, *treemap()* of **Treemap** package is used to plot a treemap showing the distribution of median unit prices and total unit sold of resale condominium by geographic hierarchy in 2017.

First, we will select records belongs to resale condominium property type from *realis2018_selected* data frame.

```{r}
realis2018_selected <- realis2018_summarised %>%
  filter(`Property Type` == "Condominium", `Type of Sale` == "Resale")
```

## 3.2 Using the basic arguments

The code chunk below designed a treemap by using three core arguments of *treemap()*, namely: *index*, *vSize* and *vColor*.

```{r}
treemap(realis2018_selected,
        index=c("Planning Region", "Planning Area", "Project Name"),
        vSize="Total Unit Sold",
        vColor="Median Unit Price ($ psm)",
        title="Resale Condominium by Planning Region and Area, 2017",
        title.legend = "Median Unit Price (S$ per sq. m)"
        )
```

Things to learn from the three arguments used:

-   index

    -   The index vector must consist of **at least two column** names or else no hierarchy treemap will be plotted.
    -   If multiple column names are provided, such as the code chunk above, the first name is the highest aggregation level, the second name the second highest aggregation level, and so on.

-   vSize

    -   The column **must not contain negative values**. This is because it’s vaues will be used to map the sizes of the rectangles of the treemaps.

*Warning:*

The treemap above was **wrongly coloured**. For a correctly designed treemap, the colours of the rectagles should be in different intensity showing, in our case, median unit prices.

## 3.3 Working with *vColor* and *type* arguments

-   The rectangles are coloured with different intensity of green, reflecting their respective median unit prices.

-   The legend reveals that the values are binned into ten bins, i.e. 0-5000, 5000-10000, etc. with an equal interval of 5000.

```{r}
treemap(realis2018_selected,
        index=c("Planning Region", "Planning Area", "Project Name"),
        vSize="Total Unit Sold",
        vColor="Median Unit Price ($ psm)",
        type = "value",
        title="Resale Condominium by Planning Region and Area, 2017",
        title.legend = "Median Unit Price (S$ per sq. m)"
        )
```

::: callout-tip
## Colours in Treemap

There are two modes: "value" vs. "manual"

-   value mode (for diverging palettes, such as "RdYlBu"):

    -   0 → intermediate color (usually white/yellow)
    -   -max(abs(values)) → left end color
    -   max(abs(values)) → right end color

-   manual mode:

    -   min(values) → left end color

    -   max(values) → right end color

    -   mean(range(values)) → middle color

Simply, value is suitable for divergent color mapping, while manual corresponds to color linearly! 🚀
:::

## 3.4 The “value” type treemap

```{r}
treemap(realis2018_selected,
        index=c("Planning Region", "Planning Area", "Project Name"),
        vSize="Total Unit Sold",
        vColor="Median Unit Price ($ psm)",
        type="value",
        palette="RdYlBu", 
        title="Resale Condominium by Planning Region and Area, 2017",
        title.legend = "Median Unit Price (S$ per sq. m)"
        )
```

-   although the colour palette used is RdYlBu but there are no red rectangles in the treemap above. This is because all the median unit prices are positive.

-   The reason why we see only 5000 to 45000 in the legend is because the *range* argument is by default c(min(values, max(values)) with some pretty rounding.

## 3.5 The “manual” type treemap

The “manual” type does not interpret the values as the “value” type does. Instead, the value range is mapped linearly to the colour palette.\
Here is the example of the inappropriate color applictaion:\
The colour scheme used is very confusing. This is because mapping = (min(values), mean(range(values)), max(values)). It is not wise to use diverging colour palette such as *RdYlBu* if the values are all positive or negative

```{r}
treemap(realis2018_selected,
        index=c("Planning Region", "Planning Area", "Project Name"),
        vSize="Total Unit Sold",
        vColor="Median Unit Price ($ psm)",
        type="manual",
        palette="RdYlBu", 
        title="Resale Condominium by Planning Region and Area, 2017",
        title.legend = "Median Unit Price (S$ per sq. m)"
        )
```

To overcome this problem, a single colour palette such as Blues should be used.

```{r}
treemap(realis2018_selected,
        index=c("Planning Region", "Planning Area", "Project Name"),
        vSize="Total Unit Sold",
        vColor="Median Unit Price ($ psm)",
        type="manual",
        palette="Blues", 
        title="Resale Condominium by Planning Region and Area, 2017",
        title.legend = "Median Unit Price (S$ per sq. m)"
        )
```

## 3.6 Treemap Layout

*treemap()* supports two popular treemap layouts, namely: “squarified” and “pivotSize”. The default is “pivotSize”.

The squarified treemap algorithm (Bruls et al., 2000) produces good aspect ratios, but ignores the sorting order of the rectangles (sortID). The ordered treemap, pivot-by-size, algorithm (Bederson et al., 2002) takes the sorting order (sortID) into account while aspect ratios are still acceptable.

## 3.7 Working with *algorithm* argument

::: panel-tabset
## squarified

✅ **Recommended**: Keeps rectangles close to square shape (default)

```{r}
treemap(realis2018_selected,
        index=c("Planning Region", "Planning Area", "Project Name"),
        vSize="Total Unit Sold",
        vColor="Median Unit Price ($ psm)",
        type="manual",
        palette="Blues", 
        algorithm = "squarified",
        title="Resale Condominium by Planning Region and Area, 2017",
        title.legend = "Median Unit Price (S$ per sq. m)"
        )
```

## pivotSize

Sorts rectangles based on size

```{r}
treemap(realis2018_selected,
        index=c("Planning Region", "Planning Area", "Project Name"),
        vSize="Total Unit Sold",
        vColor="Median Unit Price ($ psm)",
        type="manual",
        palette="Blues", 
        algorithm = "pivotSize",
        title="Resale Condominium by Planning Region and Area, 2017",
        title.legend = "Median Unit Price (S$ per sq. m)"
        )
```
:::

## 3.8 Using *sortID*

Use “pivotSize” with *sortID*  to dertemine the order in which the rectangles are placed from top left to bottom right.

```{r}
treemap(realis2018_selected,
        index=c("Planning Region", "Planning Area", "Project Name"),
        vSize="Total Unit Sold",
        vColor="Median Unit Price ($ psm)",
        type="manual",
        palette="Blues", 
        algorithm = "pivotSize",
        sortID = "Median Transacted Price",
        title="Resale Condominium by Planning Region and Area, 2017",
        title.legend = "Median Unit Price (S$ per sq. m)"
        )
```

# 4. Designing Treemap using treemapify Package

**treemapify** is a R package specially developed to draw treemaps in **ggplot2**. In this section, you will learn how to designing treemps closely resemble treemaps designing in previous section by using **treemapify**. Before you getting started, you should read [Introduction to “treemapify”](https://cran.r-project.org/web/packages/treemapify/vignettes/introduction-to-treemapify.html) its [user guide](https://cran.r-project.org/web/packages/treemapify/treemapify.pdf).

## 4.1 Designing a basic treemap

```{r}
ggplot(data=realis2018_selected, 
       aes(area = `Total Unit Sold`,
           fill = `Median Unit Price ($ psm)`),
       layout = "scol",
       start = "bottomleft") + 
  geom_treemap() +
  scale_fill_gradient(low = "light blue", high = "blue")
```

## **4.2 Defining hierarchy**

Group by Planning Region

```{r}
ggplot(data=realis2018_selected, 
       aes(area = `Total Unit Sold`,
           fill = `Median Unit Price ($ psm)`,
           subgroup = `Planning Region`),
       start = "topleft") + 
  geom_treemap()
```

Adding boundary line

```{r}
ggplot(data=realis2018_selected, 
       aes(area = `Total Unit Sold`,
           fill = `Median Unit Price ($ psm)`,
           subgroup = `Planning Region`,
           subgroup2 = `Planning Area`)) + 
  geom_treemap() +
  geom_treemap_subgroup2_border(colour = "gray40",
                                size = 2) +
  geom_treemap_subgroup_border(colour = "gray20")
```

# 5. Designing Interactive Treemap :`d3treeR`

## 5.1 Installing *d3treeR* package

1.  If this is the first time you install a package from github, you should install *devtools* package by using the code below or else you can skip this step.

```{r}
chooseCRANmirror(graphics = FALSE, ind = 1)
install.packages("devtools")
```

2.  Next, you will load the *devtools* library and install the package found in github by using the codes below.

```{r}
library(devtools)
install_github("timelyportfolio/d3treeR")
```

Now, you are ready to launch *d3treeR* package.

```{r}
library(d3treeR)
```

## 5.2 Designing An Interactive Treemap

1.  *treemap()* is used to build a treemap by using selected variables in condominium data.frame. The treemap created is save as object called ***tm***.

```{r}
tm <- treemap(realis2018_summarised,
        index=c("Planning Region", "Planning Area"),
        vSize="Total Unit Sold",
        vColor="Median Unit Price ($ psm)",
        type="value",
        title="Private Residential Property Sold, 2017",
        title.legend = "Median Unit Price (S$ per sq. m)"
        )
```

2.  Then *d3tree()* is used to build an interactive treemap.

```{r}
d3tree(tm,rootname = "Singapore" )
```

# 6. Reference

[Kam, T.S. (2025). Visual Statistical Analysis.](https://r4va.netlify.app/chap16#treemap-layout)
