---
title: "Hands on Exercise 07: Choropleth Mapping with R"
author: "Chen.PengWei"
date-modified: "last-modified"
execute:
  echo: true
  eval: true
  warning: false
  freeze: true
---

# 1. Overview

# 2.Getting Started

::: panel-tabset
## Installing Libraries

In this hands-on exercise, the key R package use is [**tmap**](https://cran.r-project.org/web/packages/tmap/) package in R. Beside **tmap** package, four other R packages will be used. They are:

-   [**readr**](https://readr.tidyverse.org/) for importing delimited text file,

-   [**tidyr**](https://tidyr.tidyverse.org/) for tidying data,

-   [**dplyr**](https://dplyr.tidyverse.org/) for wrangling data and

-   [**sf**](https://cran.r-project.org/web/packages/sf/index.html) for handling geospatial data.

Among the four packages, **readr**, **tidyr** and **dplyr** are part of **tidyverse** package.

```{r}
pacman::p_load(sf, tmap, tidyverse)
```

## Importing **Geospatial Data**

The code chunk below uses the *st_read()* function of **sf** package to import `MP14_SUBZONE_WEB_PL` shapefile into R as a simple feature data frame called `mpsz`.

```{r}
mpsz <- st_read(dsn = "data/geospatial", 
                layer = "MP14_SUBZONE_WEB_PL")
```

Only the first ten records are displayed because the dataset is being printed in a truncated format. In R, when working with **sf** (Simple Features) objects or data frames, the default behavior of the print() function is to show only the first few rows (typically 10) to prevent overwhelming the console with too much data.

```{r}
mpsz
```

## Importing **Attribute Data**

Next, we will import respopagsex2011to2020.csv file into RStudio and save the file into an R dataframe called popagsex.

```{r}

popdata <- read_csv("data/aspatial/respopagesextod2011to2020.csv")
```
:::

## 2.1 Data Preparation

Before a thematic map can be prepared, you are required to prepare a data table with year 2020 values. The data table should include the variables PA, SZ, YOUNG, ECONOMY ACTIVE, AGED, TOTAL, DEPENDENCY.

-   YOUNG: age group 0 to 4 until age groyup 20 to 24,
-   ECONOMY ACTIVE: age group 25-29 until age group 60-64,
-   AGED: age group 65 and above,
-   TOTAL: all age group, and
-   DEPENDENCY: the ratio between young and aged against economy active group
    -   ::: panel-tabset
        ## Data wrangling

        The following data wrangling and transformation functions will be used:

        -   *pivot_wider()* of **tidyr** package, and
        -   *mutate()*, *filter()*, *group_by()* and *select()* of **dplyr** package

        ```{r}
        popdata2020 <- popdata %>%
          filter(Time == 2020) %>%
          group_by(PA, SZ, AG) %>%
          summarise(`POP` = sum(`Pop`)) %>% #Computes the total population (POP) for each PA, SZ, AG group.

          ungroup() %>% #Removes grouping, allowing further transformations.

          pivot_wider(names_from=AG, 
                      values_from=POP) %>% #Converts the dataset from long format (where each row represents a single age group) to wide format, with columns for different age groups.
          mutate(YOUNG = rowSums(.[3:6])
                 +rowSums(.[12])) %>%  #Sums the population for young age groups in columns 3 to 6 and column 12.
        mutate(`ECONOMY ACTIVE` = rowSums(.[7:11])+
        rowSums(.[13:15]))%>% #sums the population for working-age individuals from columns 7 to 11 and 13 to 15
        mutate(`AGED`=rowSums(.[16:21])) %>% #All Age Groups
        mutate(`TOTAL`=rowSums(.[3:21])) %>%  
        mutate(`DEPENDENCY` = (`YOUNG` + `AGED`)
        /`ECONOMY ACTIVE`) %>%
          select(`PA`, `SZ`, `YOUNG`, 
               `ECONOMY ACTIVE`, `AGED`, 
               `TOTAL`, `DEPENDENCY`)
        ```

        ## Joining the attribute and geospatial data

        Connect the population attribute data (popdata2020) and the geographical spatial data (mpsz) to form a complete geographical attribute data frame (mpsz_pop2020), and finally save it in .rds format.

        1.  Convert the values ​​of PA and SZ fields to uppercase (uppercase)

            -   Because the SUBZONE_N and PLN_AREA_N fields in geographic data (mpsz) are both uppercase, you need to ensure that the PA and SZ of popdata2020 are also uppercase to avoid matching failures. mutate_at() is used to modify fields in batches, and toupper() converts them to uppercase.

        ```{r}
        popdata2020 <- popdata2020 %>%
          mutate_at(.vars = vars(PA, SZ), 
                  .funs = funs(toupper)) %>%
          filter(`ECONOMY ACTIVE` > 0)
        ```

        2.  Use left_join() to merge two datasets
            -   mpsz is geospatial data (sf object), containing geographic boundaries (polygons). popdata2020 is population attribute data, including population, economic activities and other information. Merged match fields
            -   mpsz is the left table, which means that even if popdata2020 has no matching data, all records of mpsz will still be retained. popdata2020 is the right table. If SZ cannot match SUBZONE_N, the population data of the district will be NA.

        ```{r}
        mpsz_pop2020 <- left_join(mpsz, popdata2020,
                                  by = c("SUBZONE_N" = "SZ"))
        ```

        3.  Stored it as .rds format

        ```{r}
        write_rds(mpsz_pop2020, "data/rds/mpszpop2020.rds")
        ```
        :::

# 3. Choropleth Mapping Geospatial Data Using *tmap*

## 3.1 Plotting a **thematic** map quickly by using *qtm()*

::: callout-tip
## Things to learn form code

tmap_mode("plot"): Set the static map mode to display the map in static map (ggplot2-like) format. tmap_mode("view"): If changed to "view", an interactive map (Leaflet-based) will be generated, which can be zoomed and clicked to view detailed information.

fill = "DEPENDENCY"： Fills the map's color based on the DEPENDENCY field data. The higher the value, the more people that need the support.
:::

```{r}
tmap_mode("plot")
qtm(mpsz_pop2020, 
    fill = "DEPENDENCY")
```

## 3.2 Creating a choropleth map by using *tmap*’s elements

*The disadvantage of qtm()* is that it makes aesthetics of individual layers harder to control. To draw a high quality cartographic choropleth map as shown in the figure below, **tmap**’s drawing elements should be used.

```{r}
tm_shape(mpsz_pop2020)+
  tm_fill("DEPENDENCY", 
          style = "quantile", 
          palette = "Blues",
          title = "Dependency ratio") +
  tm_layout(main.title = "Distribution of Dependency Ratio by planning subzone",
            main.title.position = "center",
            main.title.size = 1.2,
            legend.height = 0.45, 
            legend.width = 0.35,
            frame = TRUE) +
  tm_borders(alpha = 0.5) +
  tm_compass(type="8star", size = 2) +
  tm_scale_bar() +
  tm_grid(alpha =0.2) +
  tm_credits("Source: Planning Sub-zone boundary from Urban Redevelopment Authorithy (URA)\n and Population data from Department of Statistics DOS", 
             position = c("left", "bottom"))
```

Below is the step to draw the plot by using tmap function

:::: panel-tabset
## Drawing a base map

In the code chunk below, *tm_shape()* is used to define the input data (i.e *mpsz_pop2020*) and *tm_polygons()* is used to draw the planning subzone polygons

```{r}
tm_shape(mpsz_pop2020) +
  tm_polygons()
```

## Coloring: *`tm_polygons()`*

Assign the target variable such as *Dependency* to *tm_polygons()*.

::: callout-tip
## Things to learn form code

-   The default interval binning used to draw the choropleth map is called “pretty”.
-   The default colour scheme used is `YlOrRd` of ColorBrewer.
-   By default, Missing value will be shaded in grey.
:::

```{r}
tm_shape(mpsz_pop2020)+
  tm_polygons("DEPENDENCY")
```

#### Drawing a choropleth map using *tm_fill()* and \*tm_border()

Actually, *tm_polygons()* is a wraper of *tm_fill()* and *tm_border()*. *tm_fill()* shades the polygons by using the default colour scheme and *tm_borders()* adds the borders of the shapefile onto the choropleth map.

The *alpha* argument is used to define transparency number between 0 (totally transparent) and 1 (not transparent). By default, the alpha value of the col is used (normally 1).

Beside *alpha* argument, there are three other arguments for *tm_borders()*, they are:

-   *col* = border colour,
-   *lwd* = border line width. The default is 1, and
-   *lty* = border line type. The default is “solid”.

```{r}
tm_shape(mpsz_pop2020)+
  tm_fill("DEPENDENCY") +
  tm_borders(lwd = 0.1,  alpha = 1)
```
::::

## 3.3 Data classification methods of tmap

Most choropleth maps employ some methods of data classification. The point of classification is to take a large number of observations and group them into data ranges or classes.

**tmap** provides a total ten data classification methods, namely: *fixed*, *sd*, *equal*, *pretty* (default), *quantile*, *kmeans*, *hclust*, *bclust*, *fisher*, and *jenks*.

To define a data classification method, the *style* argument of *tm_fill()* or *tm_polygons()* will be used.

:::: panel-tabset
## jenks

Look for natural breakpoints in the data that minimize variability within each category and maximize variability between categories.

Applicable scenarios: When the data is unevenly distributed, this method can more effectively reflect the intrinsic structure of the data, such as housing prices, geographical data, etc.

```{r}
tm_shape(mpsz_pop2020)+
  tm_fill("DEPENDENCY",
          n = 5,
          style = "jenks") +
  tm_borders(alpha = 0.5)
```

## equal

Divide the data range evenly into N intervals, each interval being equal in width.

```{r}
tm_shape(mpsz_pop2020)+
  tm_fill("DEPENDENCY",
          n = 5,
          style = "equal") +
  tm_borders(alpha = 0.5)
```

::: callout-caution
Notice that the distribution of quantile data classification method are more evenly distributed then equal data classification method.
:::

## quantile

Divide the data according to its quantile (percentage), for example: -

-   quartiles
-   quintiles
-   deciles

Applicable scenarios: When the data distribution is uneven, this method ensures that the number of data in each interval is similar. For example, house price analysis, income analysis, etc.

```{r}
tm_shape(mpsz_pop2020)+
  tm_fill("DEPENDENCY",
          n = 5,
          style = "quantile") +
  tm_borders(alpha = 0.5)

```

## kmeans

Classify based on the distribution characteristics of the data itself, rather than dividing intervals evenly.

Applicable scenarios: When the data has obvious natural groupings, such as population age structure, economic status, etc.

```{r}
tm_shape(mpsz_pop2020)+
  tm_fill("DEPENDENCY",
          n = 5,
          style = "kmeans") +
  tm_borders(alpha = 0.5)
```
::::

### 3.3.1 Preparing choropleth maps based on different numbers of classes

::: panel-tabset
## 2

```{r}
tm_shape(mpsz_pop2020)+
  tm_fill("DEPENDENCY",
          n = 2,
          style = "jenks") +
  tm_borders(alpha = 0.5)
```

## 6

```{r}
tm_shape(mpsz_pop2020)+
  tm_fill("DEPENDENCY",
          n = 6,
          style = "jenks") +
  tm_borders(alpha = 0.5)
```

## 10

```{r}
tm_shape(mpsz_pop2020)+
  tm_fill("DEPENDENCY",
          n = 10,
          style = "jenks") +
  tm_borders(alpha = 0.5)
```

## 20

```{r}
tm_shape(mpsz_pop2020)+
  tm_fill("DEPENDENCY",
          n = 20,
          style = "jenks") +
  tm_borders(alpha = 0.5)
```
:::

::: callout-note
## Observations & Comparisons

**2 Classes (Low granularity)**

-   The map provides a **very generalized** view, dividing the areas into just two broad categories.

-   It may hide important variations within the dataset.

-   Useful for quick, **high-level insights**.

**6 Classes (Moderate granularity)**

-   A more **balanced** representation with clearer differentiation.

-   It provides better detail while maintaining readability.

-   Suitable for most **general analysis**.

**10 Classes (High granularity)**

-   More **detailed** but might introduce complexity.

-   The color variations may become harder to distinguish.

-   Good for **in-depth analysis**.

**20 Classes (Over-segmentation)**

-   **Too detailed** and may create excessive color variations.

-   Harder to interpret patterns.

-   Only useful when **precise differences matter**, such as in scientific analysis.
:::

### 3.3.2 Custome break point

The breakpoints can be set explicitly by means of the *breaks* argument to the *tm_fill()*. It is important to note that, in **tmap** the breaks include a minimum and maximum. As a result, in order to end up with n categories, **n+1** elements must be specified in the *breaks* option (the values must be in **increasing** order).

Now, get some descriptive statistics on the variable before setting the break points.

```{r}
summary(mpsz_pop2020$DEPENDENCY)
```

With reference to the results above, we set break point at 0.60, 0.70, 0.80, and 0.90. In addition, we also need to include a minimum and maximum, which we set at 0 and 100. Our *breaks* vector is thus c(0, 0.60, 0.70, 0.80, 0.90, 1.00).

```{r}
tm_shape(mpsz_pop2020)+
  tm_fill("DEPENDENCY",
          breaks = c(0, 0.60, 0.70, 0.80, 0.90, 1.00)) +
  tm_borders(alpha = 0.5)
```

## 3.4 Color Schceme

**tmap** supports colour ramps either defined by the user or a set of predefined colour ramps from the **RColorBrewer** package.

To change the colour, we assign the preferred colour to *palette* argument of *tm_fill()* as shown in the code chunk below.

```{r}
tm_shape(mpsz_pop2020)+
  tm_fill("DEPENDENCY",
          n = 6,
          style = "quantile",
          palette = "Blues") +
  tm_borders(alpha = 0.5)
```

To **reverse** the colour shading, add a **“-”** prefix.

```{r}
tm_shape(mpsz_pop2020)+
  tm_fill("DEPENDENCY",
          style = "quantile",
          palette = "-Greens") +
  tm_borders(alpha = 0.5)
```

## 3.5 Map Layouts

Map layout refers to the combination of all map elements into a cohensive map. Map elements include among others the objects to be mapped, the title, the scale bar, the compass, margins and aspects ratios. Colour settings and data classification methods covered in the previous section relate to the palette and break-points are used to affect how the map looks.

### 3.5.1 Map legend

::: callout-note
## Things to learn from code

-   tm_fill:

```         
-   legend.hist = TRUE: Display a histogram of the data in the legend to help understand the distribution of quantities in different categories.
-    legend.is.portrait = TRUE: Legends are arranged vertically (vertically).
-   legend.hist.z = 0.1: Set the size of the histogram, a value of 0.1 represents a smaller histogram.
```

-   tm_layout: set the map title and legend

    -   legend.outside = FALSE: The legend is placed inside the map, not outside the map.

    -   legend.position = c("right", "bottom")： The legend appears in the lower right corner.

    -   frame = FALSE: Map borders are not drawn.
:::

```{r}
tm_shape(mpsz_pop2020)+
  tm_fill("DEPENDENCY", 
          style = "jenks", 
          palette = "Blues", 
          legend.hist = TRUE, 
          legend.is.portrait = TRUE,
          legend.hist.z = 0.1) +
  tm_layout(main.title = "Distribution of Dependency Ratio by planning subzone \n(Jenks classification)",
            main.title.position = "center",
            main.title.size = 1,
            legend.height = 0.45, 
            legend.width = 0.35,
            legend.outside = FALSE,
            legend.position = c("right", "bottom"),
            frame = FALSE) +
  tm_borders(alpha = 0.5)
```

### 3.5.2 Map style

**tmap** allows a wide variety of layout settings to be changed. They can be called by using *tmap_style()*.\
 *classic* style

```{r}
tm_shape(mpsz_pop2020)+
  tm_fill("DEPENDENCY", 
          style = "quantile", 
          palette = "-Greens") +
  tm_borders(alpha = 0.5) +
  tmap_style("classic")
```

### 3.5.3 Cartographic Furniture

Beside map style, **tmap** also also provides arguments to draw other map furniture such as compass, scale bar and grid lines.

Add compass\
tm_compass() type = "8star"： Use the 8-way star compass to indicate directions on the map. size = 2: Adjust compass size.

Add grid\
tm_grid(alpha = 0.2): Add grid lines (Grid) to the map with an opacity of 0.2 to provide geographical reference points.

tm_credits() is used to indicate data sources.

```{r}
tm_shape(mpsz_pop2020)+
  tm_fill("DEPENDENCY", 
          style = "quantile", 
          palette = "Blues",
          title = "No. of persons") +
  tm_layout(main.title = "Distribution of Dependency Ratio \nby planning subzone",
            main.title.position = "center",
            main.title.size = 1.2,
            legend.height = 0.45, 
            legend.width = 0.35,
            frame = TRUE) +
  tm_borders(alpha = 0.5) +
  tm_compass(type="8star", size = 2) +
  tm_grid(lwd = 0.1, alpha = 0.2) +
  tm_credits("Source: Planning Sub-zone boundary from Urban Redevelopment Authorithy (URA)\n and Population data from Department of Statistics DOS", 
             position = c("left", "bottom"))
```

```{r}
tmap_mode("plot")

```

Can reset the default by doing this

```{r}
tmap_style("white")
```

## 3.6 Drawing Small Multiple Choropleth Maps

**Small multiple maps**, also referred to as **facet maps**, are composed of many maps arrange side-by-side, and sometimes stacked vertically. Small multiple maps enable the visualisation of how spatial relationships change with respect to another variable, such as time.

In **tmap**, small multiple maps can be plotted in three ways:

-   by assigning multiple values to at least one of the asthetic arguments,
-   by defining a group-by variable in *tm_facets()*, and
-   by creating multiple stand-alone maps with *tmap_arrange()*.

### 3.6.1 By assigning multiple values to at least one of the aesthetic arguments

tm_fill(c("YOUNG", "AGED")): Two variables are displayed here at the same time, namely: YOUNG and AGED

tmap_style() is one of the preset styles provided by tmap, you can choose from:\
"classic": classic style\
"cobalt": blue background\
"albatross": gray background\
"white": white background (suitable for reports or papers)

```{r}
tm_shape(mpsz_pop2020)+
  tm_fill(c("YOUNG", "AGED"),
          style = "equal", 
          palette = "Blues") +
  tm_layout(legend.position = c("right", "bottom")) +
  tm_borders(alpha = 0.5) +
  tmap_style("white")
```

In this example, small multiple choropleth maps are created by assigning multiple values to at least one of the aesthetic arguments.

???

```{r}
tm_shape(mpsz_pop2020)+
  tm_polygons(c("DEPENDENCY", "AGED"),
              style = c("equal", "quantile"),
              palette = list("Blues", "-Greens")) 
  tm_layout(legend.position = c("right", "bottom"))


```

### 3.6.2  Defining a group-by variable in *tm_facets()*

multiple small choropleth maps are created by using **tm_facets()**.




```{r}
tm_shape(mpsz_pop2020) +
  tm_fill("DEPENDENCY",
          style = "quantile",
          palette = "Blues",
          thres.poly = 0) + 
  tm_facets(by="REGION_N", 
            free.coords=TRUE, 
            drop.shapes=FALSE) +
  tm_layout(legend.show = FALSE,
            title.position = c("center", "center"), 
            title.size = 20) +
  tm_borders(alpha = 0.5)
```

### 3.6.3 Creating multiple stand-alone maps: *`tmap_arrange()`*

In this example, multiple small choropleth maps are created by creating multiple stand-alone maps with **tmap_arrange()**.

```{r}
youngmap <- tm_shape(mpsz_pop2020)+ 
  tm_polygons("YOUNG", 
              style = "quantile", 
              palette = "Blues")

agedmap <- tm_shape(mpsz_pop2020)+ 
  tm_polygons("AGED", 
              style = "quantile", 
              palette = "Blues")

tmap_arrange(youngmap, agedmap, asp=1, ncol=2)
```

## 3.7 Mappping Spatial Object Meeting a Selection Criterion

```{r}
tm_shape(mpsz_pop2020[mpsz_pop2020$REGION_N=="CENTRAL REGION", ])+
  tm_fill("DEPENDENCY", 
          style = "quantile", 
          palette = "Blues", 
          legend.hist = TRUE, 
          legend.is.portrait = TRUE,
          legend.hist.z = 0.1) +
  tm_layout(legend.outside = TRUE,
            legend.height = 0.45, 
            legend.width = 5.0,
            legend.position = c("right", "bottom"),
            frame = FALSE) +
  tm_borders(alpha = 0.5)
```

# 4. Reference

## 4.1 All about tmap package

-   [tmap: Thematic Maps in R](https://www.jstatsoft.org/article/view/v084i06)
-   [tmap](https://cran.r-project.org/web/packages/tmap/index.html)
-   [tmap: get started!](https://cran.r-project.org/web/packages/tmap/vignettes/tmap-getstarted.html)
-   [tmap: changes in version 2.0](https://cran.r-project.org/web/packages/tmap/vignettes/tmap-changes-v2.html)
-   [tmap: creating thematic maps in a flexible way (useR!2015)](http://von-tijn.nl/tijn/research/presentations/tmap_user2015.pdf)
-   [Exploring and presenting maps with tmap (useR!2017)](http://von-tijn.nl/tijn/research/presentations/tmap_user2017.pdf)

## 4.2 Geospatial data wrangling

-   [sf: Simple Features for R](https://cran.r-project.org/web/packages/sf/index.html)
-   [Simple Features for R: StandardizedSupport for Spatial Vector Data](https://journal.r-project.org/archive/2018/RJ-2018-009/RJ-2018-009.pdf)
-   [Reading, Writing and Converting Simple Features](https://cran.r-project.org/web/packages/sf/vignettes/sf2.html)

## 4.3 Data wrangling

-   [dplyr](https://dplyr.tidyverse.org/)
-   [Tidy data](https://cran.r-project.org/web/packages/tidyr/vignettes/tidy-data.html)
-   [tidyr: Easily Tidy Data with ‘spread()’ and ‘gather()’ Functions](https://cran.r-project.org/web/packages/tidyr/tidyr.pdf)
